Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-03-14 analysis/paper_draft3.Rmd
added Colophon and Disclosure Statement
added references to figures in Appendix; change "MCMC" to "HMC" in main text

For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero. Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus on how accurately our model can detect that bias, _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate). For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the $b_{curr}$ is not zero. Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits. For each scenario, we generated 1000 data sets and fit the model on those data sets. For each replicate, we generate a MCMC sample of 10,000 iterations with a burn-in period of 1000 iterations. For each replicate, we calculated the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.

@@

@@

@@

The choice of credible interval can affect bias detection as we define it, because the larger the credible interval one chooses, the greater the chance that zero falls within it, and vice-versa.  Here, we report simulation results for bias detection for 90% and 80% credible intervals for trials with fifteen, ten, and six strikes per side.

>>>>

For our simulation, we define bias detection as the proportion of times we identify bias when bias is present in the current trial, i.e., $b_{curr} > 0$ , based on whether credible intervals of the bias parameter exclude zero. Since we use a weakly-informative prior that pulls the estimate of $b_{curr}$ towards zero, we focus on how accurately our model can detect that bias, _not_ on how accurately the model can recover the true value of $b_{curr}$ (i.e., a traditional coverage rate). For instance, a 90\% bias detection rate means that in 900 of 1000 model fits, a given credible interval does not contain zero if the $b_{curr}$ is not zero. Put another way, if we set $b_{curr}>0$, we calculate the proportion of the 90\% credible intervals that lie to the right of zero (lower bound is positive) among the 1000 model fits. For each scenario, we generated 1000 data sets and fit the model on those data sets. For each replicate, we generated a HMC sample of 10,000 iterations with a burn-in period of 1000 iterations. For each replicate, we calculated the posterior mean and $80\%$, $90\%$, and $95\%$ highest posterior density intervals for $b_{curr}$.

@@

# Colophon

Figures were produced with ggplot2 [@ggplot2] or Tikz [@Tantau2022].

@@

# Disclosure Statement

The authors report there are no competing interests to declare.

@@

The choice of credible interval can affect bias detection as we define it, because the larger the credible interval one chooses, the greater the chance that zero falls within it, and vice-versa.  Here, we report simulation results for bias detection for 90% intervals for trials with fifteen, ten, and six strikes per side (Figures \@ref(fig:figbd9015), \@ref(fig:figbd9010), and \@ref(fig:figbd906), respectively), and the same for 80% credible intervals (Figures \@ref(fig:figbd8015), \@ref(fig:figbd8010), and \@ref(fig:figbd806)).
